PyPy's Approach to Virtual Machine Construction

### 2006-05-31 talk/dls2006/paper.tex
Some kind of translation overview graph.

The translation process starts from RPython source code and eventually
produces low-level code suitable for the target environment.  It can be
described as performing a series of step-wise transformations.  Each
step is based on control flow graph transformations and rewriting, and
on the ability to augment the program with further implementation code
written in Python and analysed with the suitable type system.

The front-end part of the translation process analyses the input
RPython program in two phases, as follows:\footnote{Note that the two
phases are intermingled in time, because type inference proceeds from
an entry point function and follows all calls, and thus only gradually
discovers (the reachable parts of) the input program.}

  [figure 0: flow graph and annotator, e.g. part of doc/image/translation.*
             then a stack of transformations
  ]
\end{verbatim}
%
\begin{enumerate}
\item We take as input RPython functions,\footnote{The input to our
       translation chain are indeed loaded runtime function objects,
       not source code or ASTs.  This enables a form of staged
       programming in which we can use unrestricted
       Python for meta-programming purposes at load time: the whole of
       the source program -- as Python program -- produces the RPython
       program input to the tool-chain as the object graph loaded in
       memory.  This includes both the relevant functions and prebuilt
       data.} and convert them to control flow graphs -- a structure
       amenable to analysis.  These flow graphs contain polymorphic
       operations only: in Python, almost all operations are
       dynamically overloaded by type.

@@

>>>>

The translation process starts from RPython source code and eventually
produces low-level code suitable for the target environment.  Its
architecture is shown in figure \ref{fig_arch}.  It can be described as
a front-end followed by a series of step-wise transformations.  Each
transformation step is based on control flow graph transformations and
rewriting, and on the ability to augment the program with further
implementation code written in Python and analysed with the suitable
type system.

The front-end part of the translation process analyses the input
RPython program in two phases, as follows:\footnote{Note that the two
phases are intermingled in time, because type inference proceeds from
an entry point function and follows all calls, and thus only gradually
discovers (the reachable parts of) the input program.}
%
\begin{enumerate}
\item We take as input RPython function objects,\footnote{The input to our
       translation chain are indeed loaded runtime function objects,
       not source code or ASTs.  This enables a form of staged
       programming in which we can use unrestricted
       Python for meta-programming purposes at load time: the whole of
       the source program -- as Python program -- produces the RPython
       program input to the tool-chain as the object graph loaded in
       memory.  This includes both the relevant functions and prebuilt
       data.} and convert them to control flow graphs -- a structure
       amenable to analysis.  These flow graphs contain polymorphic
       operations only: in Python, almost all operations are
       dynamically overloaded by type.

@@

\begin{figure}
\centering
\includegraphics[scale=0.667]{image/arch.pdf}
\caption{Overview of the translation process.}
\label{fig_arch}
\end{figure}
